On the Complexity of Best Arm Identification

### 2017-05-25 Complexity of Best Arm Identification/kl_complexity.Rmd
add matching algorithms to complexity of best-arm identification

>>>>

Note that while deriving the lower bound on the complexity measure, we have also derived the following lower bound on the failure probability for the two-armed fixed-budget bandit:

\begin{equation*}
p_t(\nu) \geq \exp \big(- \sum_{a=1}^2 \mathbb{E}_{\nu'}[N_a(t)] \KL(\nu_a', \nu_a) \big)
\end{equation*}

or, to not rely on the share of samples attributed to each arm,

\begin{equation*}
p_t(\nu) \geq \exp \big(- \max_{a=1,2} \KL(\nu_a', \nu_a) \big).
\end{equation*}

**Mention something about the fact that we were not able to do this for more than 2 arms and top-m bandits. How could it be extended to that? How could it be extended to the threshold bandit?**

As noted by Kaufmann et al. (2016), the advantage of this lower bound is that it makes the complexity in the fixed-confidence and the fixed-budget setting directly comparable. The disadvantage of course being that the result only holds for two-armed models. While the authors were not able to derive directly comparable lower bounds for the Top-$m$ case, they did derive a first lower bound in the Gaussian case with equal variances. While we would like to move beyond Gaussian distributions in the work at hand, the idea for the proof might show a way of how the ideas can be transferred to the thresholding bandit problem. We first state the result and then discuss the proof.

\begin{lemma}[Lemma 15, Kaufmann et al., 2016] \label{theorem:KaufmannEtAlLemma15}
Let $\nu$ and $\nu'$ be two bandit models such that $S^*_m(\nu) \neq S*_m(\nu')$. Then

\begin{equation*}
\max \big( \mathbb{P}_{\nu}(S \neq S^*_m(\nu)), \mathbb{P}_{\nu'}(S \neq S^*_m(\nu')) \big) \geq \frac{1}{4} \exp \Big(-\sum_{a=1}^{K} \mathbb{E}_{\nu}[N_a] \KL(\nu_a, \nu_a') \Big).
\end{equation*}
\end{lemma}

The following theorem only holds in the case of Gaussian bandit models with equal known variance.

\begin{theorem}[Theorem 17, Kaufmann et al., 2016] \label{theorem:KaufmannEtAlTheorem17}

Let $\mathcal{M}_m = \{\nu = (\nu_1, \dots, \nu_K): \nu_a = \mathcal{N}(\mu_a, \sigma^2), \mu_a \in \mathbb{R}, \mu_{[m]}\neq \mu_{[m+1]}\}.$ Furthermore, let $\nu$ be such that $\mu_1 > \dots > \mu_m > \mu_{m+1} > \dots > \mu_K$, and let

\begin{equation*}
H +(\nu) = \sum_{a=1}^m \frac{2\sigma^2}{(\mu_a - \mu_{m+1})^2}, \quad H^-(\nu) = \sum_{a=m+1}^K \frac{2\sigma^2}{(\mu_m - \mu_a)^2} \quad
and \quad
H(\nu) = H^+(\nu) + H^-(\nu).
\end{equation*}

There exists $a \in {1, \dots, m}$ and $b \in \{m+1, \dots, K\}$ such that the bandit model $\nu^{[a,b]}$ described in Figure 4 (Kaufmann et al., 2016) satisfies $H(\nu^{[a,b]}) < H(\nu)$ and is such that

\begin{equation*}
\max \big( p_t(\nu),p_t(\nu^{[a,b]})\big) \geq \frac{1}{4} \exp \big(-\frac{4t}{\bar{H}(\nu)}\big), \quad where \quad \bar{H}(\nu) = \frac{H(\nu) \min(H^+(\nu),H^-(\nu))}{H(\nu)+\min(H^+(\nu),H^-(\nu))}
\end{equation*}
\end{theorem}

Given that we are again in the Top-$m$ setting, the proof idea is similar to the proof of Theorem \ref{theorem:KaufmannEtAlTheorem4} in that we again consider problems where we switch arms around the limit seperating the two groups of arms as given implicitly by $\mu_{[m]}$ and $\mu_{[m+1]}$. The hope here being that these methods could be applied to the thresholding bandit problem where the threshold of course gives the limit between the arms above and below the threshold. As we have seen in chapter 1, this flip around the threshold was of course used in the proof of the lower bound in Theorem 1 of Locatelli et al. (2016). However, there, again, the distributions were assumed to be Gaussian. This brings the big advantage that the Kullback-Leibler divergence becomes symmetric, and the feedback can take values on all $\mathbb{R}$. And by using the complexity measure $H(\nu)$, it is also clear that there exists an arm that is undersampled with respect to its contribution to the overall complexity of the problem (or sampled exactly according to its contribution), which leads to leads to a convenient upper bound on this arms number of draws which is again employed in both the proof of Theorem \ref{theorem:KaufmannEtAlTheorem17} and in Locatelli et al. (2016). It is not clear yet, whether the same can also be done with other complexity measures.

We do not repeat the proof here for now.

See p20!

What remains is to present algorithms that are able to match the lower bounds derived so far.

#### Matching algorithm strategies

Given those lower bounds, all that is left is to find algorithms which are able to match the complexity. Sadly, not straightforward for Bernoullis.

Kaufmann et al. (2016) do not present algorithms matching the lower bound of Theorem \ref{theorem:KaufmannEtAlTheorem17}. They do consider algorithms that connect to Theorem \ref{theorem:KaufmannEtAlTheorem4} and Theorem \ref{theorem:KaufmannEtAlTheorem12}.

Given that Locatelli et al. (2016), and thus the problem we mostly consider in this thesis, is in the fixed-budget setting, we are most interested in the matching strategies for Theorem \ref{theorem:KaufmannEtAlTheorem12}. However, given that the lower bound applies only to two-armed bandits, the algorithms are of limited practical interest. Furthermore, the algorithms proposed require knowledge that is in general not available upfront.

For the example of Gaussian distributions, the knowledge of the (potentially different) variance is assumed which enables a strategy that pulls each arm proportional to its variance.

Even worse is the case of Bernoulli two-armed bandit models where the proposed algorithm requires knowledge about the true parameters of the distributions. That this limits the practical interest is also noted by the authors of Kaufmann et al. (2016). Thus, they suggest to use uniform sampling strategies in the case of two-armed Bernoulli bandits, and show that there is not much to be lost by doing so.

Thus, Kaufmann et al. (2016) join a number of other papers like Audibert et al. (2010) and Kaufmann and Kalyanakrishnan (2013) which struggle to find matching strategies in the fixed-budget setting that do not rely on knowledge of parameters or the complexity of the problem.

In contrast, Kaufmann et al. (2016) are able to show that their complexity measure derived for the fixed-confidence setting, $\mathcal{K}_C(\nu)$, can be upper-bounded using the upper bound of the sample complexity (number of pulls) given for the KL-LUCB algorithm in Kaufmann and Kalyanakrishnan (2013). Consequently, in the case of exponential family bandits, the complexity $\mathcal{K}_C(\nu)$ is bounded as given in equation (6) in Kaufmann et al. (2016):

\begin{equation*}
\sum_{a \in S^*_m} \frac{1}{K(\theta_a, \theta_{[m+1]})} + \sum_{a \notin S^*_m \frac{1}{K(\theta_a,\theta_{[m]})}} \leq \mathcal{K}_C(\nu) \leq 24 \min_{\theta \in [\theta_{[m+1]},\theta_{[m]}]} \sum_{a=1}^{K} \frac{1}{K^*(\theta_a, \theta)},
\end{equation*}

where $K(\theta, \theta') = \KL(\nu_{\theta}, \nu_{\theta'})$ for $(\theta, \theta') \in \Theta^2$ was defined as the shorthand for the Kullback-Leibler divergence for two distributions from the exponential family parameterized by $\theta$ and $\theta'$ respectively. $K^*(\theta,\theta') = K(\theta^*, \theta)$ is the Chernoff distribution between the two distributions, where $\theta^*$ is chosen so that $K(\theta^*,\theta) = K(\theta^*, \theta')$.

### 2017-05-25 Complexity of Best Arm Identification/kl_complexity.tex
add matching algorithms to complexity of best-arm identification

>>>>

Note that while deriving the lower bound on the complexity measure, we
have also derived the following lower bound on the failure probability
for the two-armed fixed-budget bandit:

\begin{equation*}
p_t(\nu) \geq \exp \big(- \sum_{a=1}^2 \mathbb{E}_{\nu'}[N_a(t)] \KL(\nu_a', \nu_a) \big)
\end{equation*}

or, to not rely on the share of samples attributed to each arm,

\begin{equation*}
p_t(\nu) \geq \exp \big(- \max_{a=1,2} \KL(\nu_a', \nu_a) \big).
\end{equation*}

\textbf{Mention something about the fact that we were not able to do
this for more than 2 arms and top-m bandits. How could it be extended to
that? How could it be extended to the threshold bandit?}

As noted by Kaufmann et al. (2016), the advantage of this lower bound is
that it makes the complexity in the fixed-confidence and the
fixed-budget setting directly comparable. The disadvantage of course
being that the result only holds for two-armed models. While the authors
were not able to derive directly comparable lower bounds for the
Top-\(m\) case, they did derive a first lower bound in the Gaussian case
with equal variances. While we would like to move beyond Gaussian
distributions in the work at hand, the idea for the proof might show a
way of how the ideas can be transferred to the thresholding bandit
problem. We first state the result and then discuss the proof.

\begin{lemma}[Lemma 15, Kaufmann et al., 2016] \label{theorem:KaufmannEtAlLemma15}
Let $\nu$ and $\nu'$ be two bandit models such that $S^*_m(\nu) \neq S*_m(\nu')$. Then

\begin{equation*}
\max \big( \mathbb{P}_{\nu}(S \neq S^*_m(\nu)), \mathbb{P}_{\nu'}(S \neq S^*_m(\nu')) \big) \geq \frac{1}{4} \exp \Big(-\sum_{a=1}^{K} \mathbb{E}_{\nu}[N_a] \KL(\nu_a, \nu_a') \Big).
\end{equation*}
\end{lemma}

The following theorem only holds in the case of Gaussian bandit models
with equal known variance.

\begin{theorem}[Theorem 17, Kaufmann et al., 2016] \label{theorem:KaufmannEtAlTheorem17}

Let $\mathcal{M}_m = \{\nu = (\nu_1, \dots, \nu_K): \nu_a = \mathcal{N}(\mu_a, \sigma^2), \mu_a \in \mathbb{R}, \mu_{[m]}\neq \mu_{[m+1]}\}.$ Furthermore, let $\nu$ be such that $\mu_1 > \dots > \mu_m > \mu_{m+1} > \dots > \mu_K$, and let

\begin{equation*}
H +(\nu) = \sum_{a=1}^m \frac{2\sigma^2}{(\mu_a - \mu_{m+1})^2}, \quad H^-(\nu) = \sum_{a=m+1}^K \frac{2\sigma^2}{(\mu_m - \mu_a)^2} \quad
and \quad
H(\nu) = H^+(\nu) + H^-(\nu).
\end{equation*}

There exists $a \in {1, \dots, m}$ and $b \in \{m+1, \dots, K\}$ such that the bandit model $\nu^{[a,b]}$ described in Figure 4 (Kaufmann et al., 2016) satisfies $H(\nu^{[a,b]}) < H(\nu)$ and is such that

\begin{equation*}
\max \big( p_t(\nu),p_t(\nu^{[a,b]})\big) \geq \frac{1}{4} \exp \big(-\frac{4t}{\bar{H}(\nu)}\big), \quad where \quad \bar{H}(\nu) = \frac{H(\nu) \min(H^+(\nu),H^-(\nu))}{H(\nu)+\min(H^+(\nu),H^-(\nu))}
\end{equation*}
\end{theorem}

Given that we are again in the Top-\(m\) setting, the proof idea is
similar to the proof of Theorem \ref{theorem:KaufmannEtAlTheorem4} in
that we again consider problems where we switch arms around the limit
seperating the two groups of arms as given implicitly by \(\mu_{[m]}\)
and \(\mu_{[m+1]}\). The hope here being that these methods could be
applied to the thresholding bandit problem where the threshold of course
gives the limit between the arms above and below the threshold. As we
have seen in chapter 1, this flip around the threshold was of course
used in the proof of the lower bound in Theorem 1 of Locatelli et al.
(2016). However, there, again, the distributions were assumed to be
Gaussian. This brings the big advantage that the Kullback-Leibler
divergence becomes symmetric, and the feedback can take values on all
\(\mathbb{R}\). And by using the complexity measure \(H(\nu)\), it is
also clear that there exists an arm that is undersampled with respect to
its contribution to the overall complexity of the problem (or sampled
exactly according to its contribution), which leads to leads to a
convenient upper bound on this arms number of draws which is again
employed in both the proof of Theorem
\ref{theorem:KaufmannEtAlTheorem17} and in Locatelli et al. (2016). It
is not clear yet, whether the same can also be done with other
complexity measures.

We do not repeat the proof here for now.

See p20!

What remains is to present algorithms that are able to match the lower
bounds derived so far.

\paragraph{Matching algorithm
strategies}\label{matching-algorithm-strategies}

Given those lower bounds, all that is left is to find algorithms which
are able to match the complexity. Sadly, not straightforward for
Bernoullis.

Kaufmann et al. (2016) do not present algorithms matching the lower
bound of Theorem \ref{theorem:KaufmannEtAlTheorem17}. They do consider
algorithms that connect to Theorem \ref{theorem:KaufmannEtAlTheorem4}
and Theorem \ref{theorem:KaufmannEtAlTheorem12}.

Given that Locatelli et al. (2016), and thus the problem we mostly
consider in this thesis, is in the fixed-budget setting, we are most
interested in the matching strategies for Theorem
\ref{theorem:KaufmannEtAlTheorem12}. However, given that the lower bound
applies only to two-armed bandits, the algorithms are of limited
practical interest. Furthermore, the algorithms proposed require
knowledge that is in general not available upfront.

For the example of Gaussian distributions, the knowledge of the
(potentially different) variance is assumed which enables a strategy
that pulls each arm proportional to its variance.

Even worse is the case of Bernoulli two-armed bandit models where the
proposed algorithm requires knowledge about the true parameters of the
distributions. That this limits the practical interest is also noted by
the authors of Kaufmann et al. (2016). Thus, they suggest to use uniform
sampling strategies in the case of two-armed Bernoulli bandits, and show
that there is not much to be lost by doing so.

Thus, Kaufmann et al. (2016) join a number of other papers like Audibert
et al. (2010) and Kaufmann and Kalyanakrishnan (2013) which struggle to
find matching strategies in the fixed-budget setting that do not rely on
knowledge of parameters or the complexity of the problem.

In contrast, Kaufmann et al. (2016) are able to show that their
complexity measure derived for the fixed-confidence setting,
\(\mathcal{K}_C(\nu)\), can be upper-bounded using the upper bound of
the sample complexity (number of pulls) given for the KL-LUCB algorithm
in Kaufmann and Kalyanakrishnan (2013). Consequently, in the case of
exponential family bandits, the complexity \(\mathcal{K}_C(\nu)\) is
bounded as given in equation (6) in Kaufmann et al. (2016):

\begin{equation*}
\sum_{a \in S^*_m} \frac{1}{K(\theta_a, \theta_{[m+1]})} + \sum_{a \notin S^*_m \frac{1}{K(\theta_a,\theta_{[m]})}} \leq \mathcal{K}_C(\nu) \leq 24 \min_{\theta \in [\theta_{[m+1]},\theta_{[m]}]} \sum_{a=1}^{K} \frac{1}{K^*(\theta_a, \theta)},
\end{equation*}

where \(K(\theta, \theta') = \KL(\nu_{\theta}, \nu_{\theta'})\) for
\((\theta, \theta') \in \Theta^2\) was defined as the shorthand for the
Kullback-Leibler divergence for two distributions from the exponential
family parameterized by \(\theta\) and \(\theta'\) respectively.
\(K^*(\theta,\theta') = K(\theta^*, \theta)\) is the Chernoff
distribution between the two distributions, where \(\theta^*\) is chosen
so that \(K(\theta^*,\theta) = K(\theta^*, \theta')\).
